# (PART) Statistics and R {-}



# R introduction

I started to work on 

## Basic R function

`Data structures` are variables with informaton stored in. R operates on these data structures. Numberic vector is a single entity consisting of a collection of numbers. 

`<-` is call assignment operator. 

```{r}
# semi-colon (‘;’) can be removed:
gene1_count <- 100;  
gene1_count
class(gene1_count)
```

```{r}
gene1_count <- c(100)
gene1_count
class(gene1_count)
```



A semi-colon (`;`) or a newline are used to separate commands 


```{r}
gene_counts <- c(5, 6, 100, 100, 200)
gene_counts
class(gene_counts)
```

```{r}
gene1_info <- c(6, "TF")
gene1_info
class(gene1_info)
```

## Producing Simple Graphs with R

The credit of this section goes to Dr. Frank McCown (@simpleGraphR). 

### Line Charts

```{r}
# Define the gene_expr_level vector with 5 values
gene_expr_level <- c(8, 20, 20, 100, 120)

# Graph the gene_expr_fpkm vector with all defaults
plot(gene_expr_level)
```


Let's add a title, a line to connect the points, and some color:

```{r}
# Define the gene_expr_level vector with 5 values
geneX_expr <- c(8, 20, 20, 100, 120)

# Graph cars using blue points overlayed by a line 
plot(geneX_expr, type="o", col="blue")

# Create a title with a red, bold/italic font
title(main="GeneX", col.main="red", font.main=4)
```

Now let's add a red line for trucks and specify the y-axis range directly so it will be large enough to fit the truck data:

```{r}
# Define the gene_expr_level vector with 5 values
geneX_expr <- c(8, 20, 20, 100, 120)
geneY_expr <- c(300, 280, 20, 10, 12)

# Graph cars using blue points overlayed by a line 
plot(geneX_expr, type="o", col="blue", ylim=c(0,300))
# Graph trucks with red dashed line and square points
lines(geneY_expr, type="o", pch=22, lty=2, col="red")
# Create a title with a red, bold/italic font
title(main="Gene expresion level", col.main="red", font.main=4)
```





## XXX


```{r}
fruit = c("apple", "apple", "pear", "orange")
fruit == "apple" 
```
```{r}
fruit = c("apple", "apple", "pear", "orange")
which(fruit == "apple")
```


```{r}
fruit = c("apple", "apple", "pear", "orange")
which(fruit == "apple" | fruit == "pear")
```


## Logic `&&` and `|`


The short answer is that && and || only ever return a single (scalar, length-1 vector) TRUE or FALSE value, whereas | and & return a vector after doing element-by-element comparisons.

The only place in R you routinely use a scalar TRUE/FALSE value is in the conditional of an if statement, so you'll often see && or || used in idioms like:

if (length(x) > 0 && any(is.na(x))) { do.something() }

In most other instances you'll be working with vectors and use & and | instead.




## List as dictionary

the list type is a good approximation. You can use names() on your list to set and retrieve the 'keys':


```{r}
foo <- vector(mode="list", length=3)
names(foo) <- c("tic", "tac", "toe")
foo[[1]] <- 12; foo[[2]] <- 22; foo[[3]] <- 33
foo
names(foo)
```



## Parsing arguments as string


### String as xlim

```{r}

```


### How to access data frame column using variable


```{r}
a = "col1"
b = "col2"
d = data.frame(a=c(1,2,3),b=c(4,5,6))
colnames(d) <- c("col1", "col2")
d[[a]]
```

This is useful when you parse a variable from the command line 


### How to create a formula from a 

It can be useful to create a formula from a string. This often occurs in functions where the formula arguments are passed in as strings.



It can be useful to create a formula from a string. This often occurs in functions where the formula arguments are passed in as strings.

```{r}
design1 = "diet"
design2 = "age"
## `~ diet + age`
as.formula(paste0("~ " , design1, " + ", design2))

```


```{r comment=''}
cat(readLines('code_R/parse_aug_as.formula.R'), sep = '\n')
```


```{sh}
/opt/R/3.5.3/lib/R/bin/Rscript code_R/parse_aug_as.formula.R time Time
```
