# Basic statistics - Descriptive Statistics 

## Probability distribution 


### Geometirc distribution

The geometric distribution is the distribution of the number of trials needed to get get the first success in repeated independent Bernolli trails. 


### Bionomial distribution 

The binomial distribution is the distribution of the number of successes (X) in  a fixed number (_n_) if independent Bernolli trails. 

### Negative binomial distribution


The negative bionomial distribution distribution is the distribution of the number of trials needed (X) to get the _r_th success (r)(@jbstatistics). 



$$
\left(\begin{array}{l}
{x-1} \\
{r-1}
\end{array}\right) p^{r-1}(1-p)^{(x-1)-(r-1)}
$$


The probabilty of the rth success that occures on the xth trial is: 

$$
\begin{aligned}
P(X=x) &=p \times\left(\begin{array}{c}
{x-1} \\
{r-1}
\end{array}\right) p^{r-1}(1-p)^{(x-1)-(r-1)} \\
&=\left(\begin{array}{c}
{x-1} \\
{r-1}
\end{array}\right) p^{r}(1-p)^{x-r}
\end{aligned}
$$

We say Random varaible X is distributed $$X \sim N B(r, p)$$. The mean is: $$\mu=\frac{r}{p}$$. The variance is $$\sigma^{2}=\frac{r(1-p)}{p^{2}}$$. 

_X_ is the count of independent Bernoulli trials required to achieve the rth successful trial when the probability of success is constant _p_.


R function `dnbinom` gives the density. `dnbbinom(x, size, prob)` calculate the probability that a number of failures _x_ occurs before _r_-th success, in a sequence of Bernoulli trials, for which the probability of individual success is p.

Example 1 (@NBblogMichael): 

An oil company has a `p = 0.20` chance of striking oil when drilling a well. What is the probability the company drills x = 7 wells to strike oil r = 3 times?

```{r}
r = 3
p = 0.20
n = 7 - r
# exact
dnbinom(x = n, size = r, prob = p)
```

```{r}
# simulated
mean(rnbinom(n = 10000, size = r, prob = p) == n)
```

```{r}
barplot(dnbinom(1:25,2,0.5), col="grey", names.arg=1:25)

```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))

data.frame(x = 0:10, prob = dnbinom(x = 0:10, size = r, prob = p)) %>%
  mutate(Failures = ifelse(x == n, n, "other")) %>%
ggplot(aes(x = factor(x), y = prob, fill = Failures)) +
  geom_col() +
  geom_text(
    aes(label = round(prob,2), y = prob + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Probability of r = 3 Successes in X = 7 Trials",
       subtitle = "NB(3,.2)",
       x = "Failed Trials (X - r)",
       y = "Probability") 
```

```{r}
#What is the expected number of trials to achieve r = 3 successes when the probability of success is p = 0.2?
r = 3
p = 0.20
# mean
# exact
r / p

```
```{r}
# simulated
mean(rnbinom(n = 10000, size = 3, prob = p)) + r
```

```{r}
# Variance
# exact
r * (1 - p) / p^2
```
```{r}
# simulated
var(rnbinom(n = 100000, size = r, prob = p))
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))

data.frame(x = 1:20, 
           pmf = dnbinom(x = 1:20, size = r, prob = p),
           cdf = pnbinom(q = 1:20, size = r, prob = p, lower.tail = TRUE)) %>%
ggplot(aes(x = factor(x), y = cdf)) +
  geom_col() +
  geom_text(
    aes(label = round(cdf,2), y = cdf + 0.01),
    position = position_dodge(0.9),
    size = 3,
    vjust = 0
  ) +
  labs(title = "Cumulative Probability of X = x failed trials to achieve 3rd success",
       subtitle = "NB(3,.2)",
       x = "Failed Trials (x)",
       y = "probability") 

```

#### Why is it called Negative Binomial?



The “negative” part of negative binomial stems from the fact that one facet of the binomial distribution is reversed: in a binomial experiment, we count the number of successes in a fixed number of trials. In a negative binomial experiment, we’re counting the failures, or how many cards it takes you to pick two aces.


